---
title: "MICROBIOME MASTER ITS (v.0.2)"
date: "`r Sys.Date()`"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(options(scipen=999))
library(ggplot2)
library(tidyverse)
library(data.table)
library(reshape2)
library(RColorBrewer)
library(viridis)
#library(pastecs)
```

```{r run_analyses, include=FALSE}
DATA <- read.table("../../results/qiime-dada2-module/feature_table/denoising_stats/stats.tsv", sep="\t", header=TRUE, comment.char="#")
DATA1 <-  reshape2::melt(DATA, id.vars=c("sample.id"), 
                         measure.vars=c("input",
                                        "filtered",
                                        "percentage.of.input.passed.filter",
                                        "denoised",
                                        "merged",
                                        "percentage.of.input.merged",
                                        "non.chimeric",
                                        "percentage.of.input.non.chimeric"))

PRCT      <- c("percentage.of.input.passed.filter", "percentage.of.input.merged", "percentage.of.input.non.chimeric")
NUM       <- c("input", "denoised", "merged", "non.chimeric")
DATA_PRCT <- filter(DATA1, variable %in% PRCT)
DATA_NUM  <- filter(DATA1, variable %in% NUM)

# Get important stat values
SAMPLE_NO <- nrow(DATA)

INPUT_MEAN <- round(mean(DATA$input),2)
INPUT_STD <- round(sd(DATA$input),2)
INPUT_MIN <-min(DATA$input)
INPUT_MAX <- max(DATA$input)

DENOISED_MEAN <- round(mean(DATA$denoised),2)
DENOISED_STD <- round(sd(DATA$denoised),2)
DENOISED_MIN <-min(DATA$denoised)
DENOISED_MAX <- max(DATA$denoised)

MERGED_MEAN <- round(mean(DATA$merged),2)
MERGED_STD <- round(sd(DATA$merged),2)
MERGED_MIN <-min(DATA$merged)
MERGED_MAX <- max(DATA$merged)

NC_MEAN <- round(mean(DATA$non.chimeric),2)
NC_STD <- round(sd(DATA$non.chimeric),2)
NC_MIN <-min(DATA$non.chimeric)
NC_MAX <- max(DATA$non.chimeric)

generate_barplots <- function(d,l) {
  plot <- ggplot(d, aes(x=reorder(sample.id, -value), y=value, fill=variable)) +
    geom_bar(position="stack", stat="identity") +
    scale_y_continuous(labels = scales::comma) +
    scale_fill_viridis(discrete = TRUE) +
    xlab("Sample/library") +
    ylab(paste(l)) +
    theme(legend.position = "top") + 
    theme(axis.text.x=element_blank(),
          axis.ticks.x=element_blank())
  
  return(plot)
  
}

generate_histograms <- function(d,l,t,b) {
  plot <- ggplot(d, aes(x=.data[[l]])) +
                   geom_histogram(binwidth=b, fill="#69b3a2", color="#e9ecef", alpha=0.9) +
                   scale_y_continuous(labels = scales::comma) +
                   scale_x_continuous(labels = scales::comma) +
                   xlab(paste(t)) +
                   ylab("Count") 
  return(plot)
  
}

generate_boxplots <- function(d,l) {
  plot <- ggplot(d, aes(variable, value, colour=variable)) +
    geom_point(alpha=0.01) +
    geom_jitter(position=position_jitter(width=.2, height=0)) +
    geom_boxplot(outlier.shape=NA, alpha=0.4,lwd=0.75) +
    scale_y_continuous(labels = scales::comma) +
    scale_colour_viridis(discrete = TRUE) +    
    ylab(paste(l)) +
    theme(legend.position = "right") +
    theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())
  
  return(plot)
  
}

```

## DADA2 OUTPUT REPORT

---

### General Statistics

**Sample No**: `r SAMPLE_NO`

**Input**  
Min/Max: `r INPUT_MIN` / `r INPUT_MAX`  
Average: `r INPUT_MEAN` ± `r INPUT_STD`  

**Denoised**  
Min/Max: `r DENOISED_MIN` / `r DENOISED_MAX`  
Average: `r DENOISED_MEAN` ± `r DENOISED_STD`  

**Merged**  
Min/Max: `r MERGED_MIN` / `r MERGED_MAX`  
Average: `r MERGED_MEAN` ± `r MERGED_STD`  

**Non-chimeric**  
Min/Max: `r NC_MIN` / `r NC_MAX`  
Average: `r NC_MEAN` ± `r NC_STD`  

---

### Summary Plots

**Histogram of read counts AFTER denoising** 


```{r print_hist1, include=TRUE, echo=FALSE, fig.align='left'}
generate_histograms(DATA, "non.chimeric", "Number of reads after denoising processing", 5000)
```


**Histogram of % of reads AFTER denoising steps** 


```{r print_hist2, include=TRUE, echo=FALSE, fig.align='left'}
generate_histograms(DATA, "percentage.of.input.non.chimeric", "% of reads after denoising processing", 5)
```


**Number of reads per sample at denoising steps** 


```{r print_barplots1, include=TRUE, echo=FALSE, fig.align='left'}
generate_barplots(DATA_NUM,"Number of reads")
```


**% of reads per sample after denoising steps** 


```{r print_barplots2, include=TRUE, echo=FALSE, fig.align='left'}
generate_barplots(DATA_PRCT, "% of reads")
```


**Number of Reads AFTER denoising** 

```{r print_boxplots2, include=TRUE, echo=FALSE, fig.align='left'}
generate_boxplots(DATA_NUM, "Number of reads after denoising")
```


**% of Reads AFTER denoising** 

```{r print_boxplots1, include=TRUE, echo=FALSE, fig.align='left'}
generate_boxplots(DATA_PRCT,"% of reads after denoising")
```



*--- End of Report ---*